---
title: "Assignment 3"
author:
- Zekai Li
- S2040608
header-includes:
- \usepackage{amssymb}
- \usepackage{amsmath}
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies:
    - amssymb
    - amsmath
---

```{r setup, include=FALSE}
tinytex::install_tinytex()
knitr::opts_chunk$set(echo = TRUE)
```

```{r library packages, echo=FALSE, results='hide', include=FALSE}
library(purrr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(maxLik)
library(mice)
```

## Question 1

### $(a)$ My solution
```{r}
print(c("The propotion of incomplete cases is",
        1-nrow(cc(nhanes))/nrow(nhanes)),quote=FALSE)
```
The percentage of incomplete cases is 48%.

### $(b)$ My solution
```{r}
#nhanes
imps <- mice(nhanes,method="norm",printFlag=FALSE,seed=1)
fits <- with(imps,lm(bmi~age+hyp+chl))
ests <- pool(fits)
print(ests[,3][,c(3,4,5,9,10)])
#summary(ests,conf.int=TRUE)[,c(2,3,6,7,8)]
```
From the pooled estimates, the proportions of variance due to missing data($\frac{B+\frac{B}{M}}{V^\top}$) for intercept is 0.2233797, for the coefficient of "age" is 0.5997594, for the coefficient of "hyp" is 0.4857348, for the coefficient of "chl" is 0.4220811. Looking at the riv, *relative increase in variance*($\frac{B+\frac{B}{M}}{\bar{U}}$). "riv" of the coefficient of "age" is the largest. Therefore, the parameter for "age" appear to be most affected by the nonresponse.

### $(c)$ My solution
```{r}
ests_2 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=2),lm(bmi~age+hyp+chl)))
ests_3 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=3),lm(bmi~age+hyp+chl)))
ests_4 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=4),lm(bmi~age+hyp+chl)))
ests_5 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=5),lm(bmi~age+hyp+chl)))
ests_6 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=6),lm(bmi~age+hyp+chl)))
ests_2[,3][,c(3,4,5,9,10)]
ests_3[,3][,c(3,4,5,9,10)]
ests_4[,3][,c(3,4,5,9,10)]
ests_5[,3][,c(3,4,5,9,10)]
ests_6[,3][,c(3,4,5,9,10)]
```

### $(d)$ My solution
```{r}
ests_m1 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=1,m=100),lm(bmi~age+hyp+chl)))
ests_m2 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=2,m=100),lm(bmi~age+hyp+chl)))
ests_m3 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=3,m=100),lm(bmi~age+hyp+chl)))
ests_m4 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=4,m=100),lm(bmi~age+hyp+chl)))
ests_m5 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=5,m=100),lm(bmi~age+hyp+chl)))
ests_m6 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=6,m=100),lm(bmi~age+hyp+chl)))
ests_m1[,3][,c(3,4,5,10)]
ests_m2[,3][,c(3,4,5,10)]
ests_m3[,3][,c(3,4,5,10)]
ests_m4[,3][,c(3,4,5,10)]
ests_m5[,3][,c(3,4,5,10)]
ests_m6[,3][,c(3,4,5,10)]
```

## Question 2

### My solution
```{r}
load("dataex2.Rdata")

# built a function to return a matrix with a row 95% confidence intervals for beta_1
# input:  ## D, 2-D, from dataex2
          ## Mtd, a string used in mice() function
# output: ## a matrix
CI_beta1 <- function(D,Mtd){
  imps <- mice(data=D,method=Mtd,printFlag=FALSE,seed=1,m=20)
  ests <- pool(with(imps,lm(Y~X)))
  return(matrix(summary(ests,conf.int=TRUE)[,c(7,8)][2,],nrow=1))
}

# built a function to return a matrix with rows 95% confidence intervals for beta_1
# input:  ## data, 3-D,initialized as dataex2
          ## Mtd, a string used in mice() function
# output: ## a matrix
CIs <- function(data=dataex2,Mtd){
  Conf <- matrix(nrow=1,ncol=2)
  for (i in 1:100){
    A <- CI_beta1(data[,,i],Mtd=Mtd)
    Conf <- rbind(Conf,A)
  }
  return(Conf[2:101,])
}

# function to return a emprical coverage probability
  ## with 95% confidence intervals for beta_1
# input:  ## data, 3-D,initialized as dataex2
          ## Mtd, a string used in mice() function
# output: ## emprical coverage probability(numeric)
ECP <- function(data=dataex2,Mtd){
  n = 100; times = 0
  B = CIs(Mtd=Mtd)
  for (i in 1:100){
    times = ifelse(B[i,][1]<3 & B[i,][2]>3,times+1,times)
  }
  return(times/n)
}

print(c("The emprical probability for ß1 using stochastic regression imputation is",ECP(Mtd="norm.nob")),quote=FALSE)
print(c("The emprical probability for ß1 using the corresponding bootstrap based version is",ECP(Mtd="norm.boot")),quote=FALSE)
```

## Question 3

### My solution


## Question 4

### $(a)$ My solution
```{r}
load("dataex4.Rdata")
dataex4
ests <- pool(with(mice(dataex4,printFlag=FALSE,seed=1,m=50),lm(y~x1*x2)))
summary(ests,conf.int=TRUE)[,c(1,2,7,8)]
```
The estimates for $\beta_1$ is 1.4112333, and the 95% confident interval is $[1.219397,1.6030697]$;
The estimates for $\beta_1$ is 1.9658191, and the 95% confident interval is $[1.860657,2.0709812]$;
The estimates for $\beta_1$ is 0.7550367, and the 95% confident interval is $[0.642302,0.8677715]$.

### $(b)$ My solution


### $(c)$ My solution


### $(d)$ My solution


## Question 5




