---
title: "Assignment 3"
author:
- Zekai Li
- S2040608
header-includes:
- \usepackage{amssymb}
- \usepackage{amsmath}
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies:
    - amssymb
    - amsmath
---

```{r setup, include=FALSE}
tinytex::install_tinytex()
knitr::opts_chunk$set(echo = TRUE)
```

```{r library packages, echo=FALSE, results='hide', include=FALSE}
library(purrr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(maxLik)
library(mice)
```

## Question 1

### $(a)$ My solution
```{r}
print(c("The propotion of incomplete cases is",
        1-nrow(cc(nhanes))/nrow(nhanes)),quote=FALSE)
```
The percentage of incomplete cases is 48%.

### $(b)$ My solution
```{r}
#nhanes
imps <- mice(nhanes,method="norm",printFlag=FALSE,seed=1)
fits <- with(imps,lm(bmi~age+hyp+chl))
ests <- pool(fits)
ests
summary(ests,conf.int=TRUE)[,c(2,3,7,8)]
```
From the pooled estimates, the proportions of variance due to missing data for intercept is 0.2233797, for the coefficient of "age" is 0.5997594, for the coefficient of "hyp" is 0.4857348, for the coefficient of "chl" is 0.4220811

### $(c)$ My solution
```{r}
ests_2 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=2),lm(bmi~age+hyp+chl)))
ests_3 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=3),lm(bmi~age+hyp+chl)))
ests_4 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=4),lm(bmi~age+hyp+chl)))
ests_5 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=5),lm(bmi~age+hyp+chl)))
ests_6 <- pool(with(mice(nhanes,method="norm",printFlag=FALSE,seed=6),lm(bmi~age+hyp+chl)))
summary(ests_3)
```

### $(d)$ My solution


## Question 2


## Question 3

### $(i)$ My solution


### $(ii)$ My solution


## Question 4

### $(a)$ My solution


### $(b)$ My solution


### $(c)$ My solution


### $(d)$ My solution


## Question 5




